---
title: "p8105_hw2_jc4166"
author: "Jerri Chen"
date: "2019_09_23"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

The "Mr. Trash Wheel" sheet was read in from the "HealthyHarborWaterWheelTotals2018-7-28.xlsx" Excel file and variable names were cleaned. Rows and columns that did not contain dumpster-specific data were omitted, and for the variable "sports_balls" the number was rounded to the nearest integer, and then converted to an integer variable.
```{r eval = FALSE}
library(readxl)
mr_trash_wheel_data = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>%
  subset(dumpster != "NA") %>%
  subset(select = -c(x15)) %>%
  mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))
```

Precipitation data for 2017 and 2018 were read and cleaned, omitting rows without precipitation data and adding a variable "year". Then, the 2017 and 2018 precipitation datasets were combined, and "month" was converted to a character variable.
```{r}
library(readxl)
precipitation_2017_data = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  subset(total != "NA") %>%
  mutate(year = "2017")

precipitation_2018_data = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  subset(total != "NA") %>%
  mutate(year = "2018")

precipitation_17_18_combined_data = 
  left_join(precipitation_2017_data, precipitation_2018_data, by = "month") %>%
  mutate(month = month.name[month])
```

The "precipitation_17_18_combined_data" dataset summarizes precipitation in inches for each month of 2017 and 2018. The data for 2017 is complete as there are entries for each month, (`r count(precipitation_2017_data, total != "NA")` observations total). The data for 2018 is not complete as several months are missing; there are `r count(precipitation_2018_data, total != "NA")` observations total. The total precipitation for available months in 2018 is XXXX inches. The "mr_trash_wheel_data" summarizes the types of objects that are collected by Mr. Trash Wheel. For example, in 2018, XXXX sports balls were collected in total.

## Problem 2
Data from "pols-month.csv" was read in and cleaned.

```{r eval = FALSE}
pols_month_data = read_csv(file = "./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(president = if_else(prez_gop == 1, "gop", "dem")) %>% 
  select(-prez_gop, -prez_dem, -day)
```

Data from "snp.csv" was read in and cleaned.
```{r eval = FALSE}
snp_data = read_csv(file = "./data/snp.csv") %>% 
  janitor::clean_names() %>%
  separate(col = date, into = c("month", "day", "year"), sep = "/") %>% 
  select("year", "month", "day", "close") %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  select(-day) %>% 
  rename(snp_close = close)
```

Data from "unemployment.csv" was read in and cleaned.
```{r eval = FALSE}
unemployment_data = read_csv(file = "./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  mutate(year = as.character(year))

unemployment_data_tidy = 
  pivot_longer(
    unemployment_data,
    jan:dec,
    names_to = "month",
    values_to = "percent_unempl") %>% 
  mutate(month = case_when(
    month == "jan" ~ "January",
    month == "feb" ~ "February",
    month == "mar" ~ "March",
    month == "apr" ~ "April",
    month == "may" ~ "May",
    month == "jun" ~ "June",
    month == "jul" ~ "July",
    month == "aug" ~ "August",
    month == "sep" ~ "September",
    month == "oct" ~ "October",
    month == "nov" ~ "November",
    month == "dec" ~ "December"
  ))
```

Data from snp was merged into pols to form pols_snp_data, and then unemployment was merged into pols_snp_data to form pols_snp_unemployment_data.

```{r eval = FALSE}
pols_snp_data = 
  left_join(pols_month_data, snp_data, by = c("year", "month"))

pols_snp_unemployment_data = 
  left_join(pols_snp_data, unemployment_data_tidy, by = c("year", "month"))
```

The pols_snp_unemployment_data dataset contains merged data from the pols_month, snp, and unemployment datasets. Each of the datasets had information delineated by dates, which was cleaned into "year" and "month" variables to be able to merge the datasets. pols_month had counts for the number of national politicians (governors, senators, representatives) of each party (gop or democrat) for each year and month from January 1947 to June 2015, and whether the president was gop or democrat. The snp dataset has the closing values of the S&P stock index by date from January 1950 to July 2015. The unemployment dataset has the percent unemployment for each month and year from January 1948 to December 2015. The resulting merged dataset (pols_snp_unemployment_data) is a tibble with 11 variables and 822 observations, though there is missing snp data from 1947 to 1949 and missing unemployment data from 1947. The data is otherwise complete until June 2015.

## Problem 3

Data from Popular_Baby_Names.csv was read in and cleaned.
```{r eval = FALSE}
baby_names_tidy = read.csv(file = "./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  rename(name = child_s_first_name) %>% 
  rename(year = year_of_birth) %>% 
  mutate(name = tolower(name)) %>% 
  mutate(gender = tolower(gender)) %>% 
  mutate(ethnicity = tolower(ethnicity)) %>% 
  mutate(ethnicity = case_when(
    ethnicity == "asian and pacific islander" ~ "asian and pacific islander",
    ethnicity == "asian and paci" ~ "asian and pacific islander",
    ethnicity == "black non hispanic" ~ "black non hispanic",
    ethnicity == "black non hisp" ~ "black non hispanic",
    ethnicity == "white non hispanic" ~ "white non hispanic",
    ethnicity == "white non hisp" ~ "white non hispanic"
  ))
```

Create a table showing the rank in popularity of the name "Olivia" over time.

