---
title: "p8105_hw2_jc4166"
author: "Jerri Chen"
date: "2019_09_23"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

The "Mr. Trash Wheel" sheet was read in from the "HealthyHarborWaterWheelTotals2018-7-28.xlsx" Excel file and variable names were cleaned. Rows and columns that did not contain dumpster-specific data were omitted, and for the variable "sports_balls" the number was rounded to the nearest integer, and then converted to an integer variable.
```{r}
library(readxl)
mr_trash_wheel_data = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>%
  subset(dumpster != "NA") %>%
  subset(select = -c(x15))
```

Precipitation data for 2017 and 2018 were read and cleaned, omitting rows without precipitation data and adding a variable "year". Then, the 2017 and 2018 precipitation datasets were then combined, and "month" was converted to a character variable.
```{r}
library(readxl)
precipitation_2017_data = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  subset(total != "NA") %>%
  mutate(year = "2017")

precipitation_2018_data = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  subset(total != "NA") %>%
  mutate(year = "2018")

precipitation_17_18_combined_data = 
  left_join(precipitation_2017_data, precipitation_2018_data, by = "month")
```
